### 2021-07-10 submission.tex
latex set up

\documentclass{article}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{parskip}
\usepackage{graphicx}

\usepackage[top=2.5cm, left=3cm, right=3cm, bottom=4.0cm]{geometry}

\newcommand{\lectureheader}[4]{%
  \begin{minipage}{.3\textwidth}%

    \strut\includegraphics[width=\textwidth]{figures/ethlogo.pdf}%
  \end{minipage} \hfill%
  \raisebox{1.5mm}{%
    \begin{minipage}{0.69\textwidth}\sf\flushright%
        \textbf{\Huge #3}\mbox{\hspace{2mm}}\\#4\mbox{\hspace{2mm}}%
    \end{minipage}%
  }\\[-2mm]\hrule%
  \begin{minipage}[t]{0.5\textwidth}\sf\textit{#1} \end{minipage} \hfill%
  \begin{minipage}[t]{0.5\textwidth}\sf\flushright \textit{#2}\end{minipage}%
  \par%
}

% Create commands for syntax that you will frequently use
\newcommand{\xx}{\mathbf x}

\begin{document}
\begin{titlepage}

\lectureheader{Prof. Ryan Cotterell}
{}
{\Large Natural Language Processing}{Spring 2021}
	\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}

	\center % Centre everything on the page
	{\Huge Course Assignment}\\
      \quad\newline

	{\large\today} \\
	\quad\newline
	%	Author
	%------------------------------------------------

	{\Large Karol Borkowski \\ \emph{nethz} Username: kborkowski \\ Student ID: 00000000}\\[0.5cm] 
	\vfill
	{\large \textbf{Collaborators:} \\
	Other student 1 \\
	Other student 2}

	\vfill\vfill\vfill 
	By submitting this work, I verify that it is my own. That is, I have written my own solutions to each problem for which I am submitting an answer. I have listed above all others with whom I have discussed these answers.

	\vfill

\end{titlepage}

\part{Course Assignment Episode 1}

\section*{Question 1} 
\begin{enumerate}[label = (\alph*)]
    \item

\begin{figure}[h!]
    \centering
    \includegraphics[scale=1.5]{figures/backprop.jpg}
    \caption{The computation graph of f.}
    \label{fig:roller-coaster}
\end{figure}

$a_1 = x_1 \cdot w_{11}^1 \; \;$
$b_1 = x_2 \cdot w_{21}^1 \; \;$
$c_1 = x_3 \cdot w_{31}^1 \; \;$

$d_1=a_1 + b_1 + c_1$

$h_1 = ReLu(d_1)$

$a_2 = x_1 \cdot w_{12}^1 \; \;$
$b_2 = x_2 \cdot w_{22}^1 \; \;$
$c_2 = x_3 \cdot w_{32}^1 \; \;$

$d_2=a_2 + b_2 + c_2$

$h_2 = ReLu(d_2)$

$a_3 = h_1 \cdot w_{11}^2 \; \;$
$b_3 = h_2 \cdot w_{21}^2 \; \;$

$d_3=a_3 + b_3$

$y=\sigma(d_3)$

\item 
	\begin{enumerate}[label = (\roman*)]
		\item
		$h_1=ReLu(1+1+1) = 3 \; \;$
		$h_2=ReLu(1+1+1) = 3 $

		$y=\sigma(3+3) \simeq 0.99753$

		\item
		$\frac{\partial L}{\partial \hat{y}} = \frac{-y}{\hat{y}}-\frac{1-y}{1-\hat{y}}\simeq -404 \; \; \;$

		$\frac{\partial y}{\partial d_3} = \sigma(d_3) \cdot (1-\sigma(d_3)) \simeq 0.00247 \; \; \;$

$\frac{\partial d_3}{\partial a_3} = 1 $

$\frac{\partial d_3}{\partial b_3} = 1 \; \; \;$

$\frac{\partial a_3}{\partial w_{11}^2} = h_1 = 3 \; \; \;$

$\frac{\partial b_3}{\partial w_{21}^2} = h_2 = 3 \; \; \;$

$\frac{\partial a_3}{\partial h_1} = w_{11}^2 = 1 $

$\frac{\partial b_3}{\partial h_2} = w_{21}^2 = 1 $

$\frac{\partial h_1}{\partial d_1} = 1 $

$\frac{\partial h_2}{\partial d_2} = 1 $

$\frac{\partial d_1}{\partial a_1} = 1 $

$\frac{\partial d_1}{\partial b_1} = 1 $

$\frac{\partial d_1}{\partial c_1} = 1 $

$\frac{\partial d_2}{\partial a_2} = 1 $

$\frac{\partial d_2}{\partial b_2} = 1 $

$\frac{\partial d_2}{\partial c_2} = 1 $

$\frac{\partial a_1}{\partial w_{11}^1} = 1 $

$\frac{\partial b_1}{\partial w_{21}^1} = 1 $

$\frac{\partial c_1}{\partial w_{31}^1} = 1 $

$\frac{\partial a_2}{\partial w_{12}^1} = 1 $

$\frac{\partial b_2}{\partial w_{22}^1} = 1 $

$\frac{\partial c_2}{\partial w_{32}^1} = 1 $

$\frac{\partial a_1}{\partial x_1} = 1 $

$\frac{\partial b_1}{\partial x_2} = 1 $

$\frac{\partial c_1}{\partial x_3} = 1 $

$\frac{\partial a_2}{\partial x_1} = 1 $

$\frac{\partial b_2}{\partial x_2} = 1 $

$\frac{\partial c_2}{\partial x_3} = 1 $

		\item
		$\frac{\partial L}{\partial \hat{y}} = \frac{-y}{\hat{y}}-\frac{1-y}{1-\hat{y}}\simeq -404 \; \; \;$

		\item
		$\Delta w_{ij} = -\eta \frac{\partial L}{\partial w_{ij}}$

		$\frac{\partial L}{\partial w_{11}^2} = \frac{\partial L}{\partial \hat{y}} \cdot \frac{\partial \hat{y}}{\partial d_3}  \cdot \frac{\partial d_3}{\partial w_{11}^2} = -0.99788$

$\Delta w_{11}^2 = 0.1 \cdot -0.99788 = -0.099788$

		$\frac{\partial L}{\partial w_{21}^2} = \frac{\partial L}{\partial \hat{y}} \cdot \frac{\partial \hat{y}}{\partial d_3}  \cdot \frac{\partial d_3}{\partial w_{21}^2} = -0.99788$

$\Delta w_{21}^2 = 0.1 \cdot -0.99788 = -0.099788$

		$\frac{\partial L}{\partial w_{11}^1} = \frac{\partial L}{\partial \hat{y}} \cdot \frac{\partial \hat{y}}{\partial d_3}  \cdot \frac{\partial d_3}{\partial a_3} \cdot \frac{\partial a_3}{\partial h_1 } \cdot \frac{\partial h_1}{\partial d_1} \cdot \frac{\partial d_1}{\partial a_1} \cdot \frac{\partial a_1}{\partial w_{11}^1} = -0.99788$

$\Delta w_{11}^1 = 0.1 \cdot -0.99788 = -0.099788$

Since the input values are the same and the weight values are also the same, the rest of the gradient changes have the same value. The calculation are analogous to the above ones.

	\end{enumerate}

	\item
	The symmetry of weights is not broken, so the hidden units can be replaced with a single unit producing the same output. The benefits of such solution are: faster training, more reliable model due to fewer parameters.

\end{enumerate}

\section*{Question 4} 
\begin{enumerate}[label = (\alph*)]
    \item
    \begin{enumerate}[label = (\roman*)]
    \item

    \item

	\end{enumerate}

	\item
    \begin{enumerate}[label = (\roman*)]
    \item

    \item

	\end{enumerate}

	\item
    \begin{enumerate}[label = (\roman*)]
    \item

    \item

    \item

    \item

	\end{enumerate}

\end{enumerate}

\clearpage
\newpage

\begin{table}[h]
        \centering
        \fontsize{10}{10}\selectfont
        \renewcommand{\arraystretch}{1.2} % vertical padding
        \setlength{\tabcolsep}{0.5em} % for the horizontal padding
        \begin{tabular}{l|l|l|l}
        \textbf{number} & \textbf{sample strings} & \textbf{accepted} & \textbf{weight} \\ \hline
        1 & educational is this not &  &  \\
        2 & is this assignment educational &  &  \\
        3 & not educational is not educational &  &  \\
        4 & this assignment is not educational &  &  \\
        5 & is this assignment educational &  &  \\
        6 & this assignment course is educational &  &  \\
        7 & is this assignment not educational &  &  \\
        8 & this assignment not &  &  \\
        9 & this course assignment is not educational &  &  \\
        10 & this course is not not educational &  &  \\
        11 & not educational is this &  &  \\
        12 & course assignment is not educational &  &  \\
        13 & not this assignment is educational &  &  \\
        14 & not not not educational &  &  \\
        14 & is this course assignment not educational &  &  \\
        15 & course assignment is this &  &  \\
        16 & this course is interesting &  &  \\
        17 & this course assignment not educational &  & 
        \end{tabular}
        \caption{Some strings from $\mathcal{Y}_{\geq 2, \leq 6}$}
        \label{tab:wfst_strings}
        \end{table}

\begin{figure}[!h]
        \centering
        \includegraphics[width=1.0\textwidth]{./figures/wfst_charts_1.pdf}
        % \includegraphics[width=.45\linewidth]{./fig/C1000.png}
        \caption{Floyd-Warshall algorithm, iteration 0 to 3; left column matrix should contain weights after iteration n; right column matrix should be iteratively filled for backtracking each path}
        \label{fig:wfst_charts_1}
    \end{figure}

    \begin{figure}[!h]
        \centering
        \includegraphics[width=1.0\textwidth]{./figures/wfst_charts_2.pdf}
        % \includegraphics[width=.45\linewidth]{./fig/C1000.png}
        \caption{Floyd-Warshall algorithm, iteration 4 to 7; left column matrix should contain weights after iteration n; right column matrix should be iteratively filled for backtracking each path}
        \label{fig:wfst_charts_2}
    \end{figure} 
\end{document}